---
title: "BestCancer"
author: "Ashley"
date: '2022-04-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library
```{r}
library("tidyverse")
library("janitor")
library("hrbrthemes")
library("psych")
library("ggrepel")
library("ggplot2")
library("ggpubr")
library("broom")
library("AICcmodavg")
library("ctv") 
library("ape")
library("phytools")
library("phangorn")
library("geiger")
library("nlme")
library("visreg")
library("tidytree")
library("ggthemes")
library("wesanderson")
library("kableExtra")
```

Data
```{r}
lh <- read_csv("Pantheria_og.csv") #life history date
pi <- read.csv("placenta__more_interdigitation.csv") # placenta data
np <- read.csv("neoplasiaPrevalence.min10.2022-04-01T19_13.csv") # cancer data
mammals <- read.newick("mammals_species.nwk") # phylogeny data
#brownian <- read.csv("brownian_pvalue.csv")%>% 
 # rename(OutcomeMeasure = "ï..Outcome.Measure")%>%
  #rename(Invasiveness.Interdigitation = "Invasiveness...Interdigitation")

```

# Wrangling Data
```{r}
# Selecting & renaming variables
lh <- lh %>%
  clean_names()

# select variables of interest, rename variables, & convert from months to days. Note to used days throughout!!
lh2 <- lh %>%
  select(msw93_binomial, x5_1_adult_body_mass_g, x3_1_ageat_first_birth_d, x9_1_gestation_len_d, x14_1_inter_birth_interval_d, x15_1_litter_size, x16_1_litters_per_year, x17_1_max_longevity_m, x5_3_neonate_body_mass_g,x23_1_sexual_maturity_age_d)%>%
  rename(species = "msw93_binomial") %>%
  rename(adultmass = "x5_1_adult_body_mass_g")%>%
  rename(agefb = "x3_1_ageat_first_birth_d") %>%
  rename(gestation = "x9_1_gestation_len_d") %>%
  rename(interv = "x14_1_inter_birth_interval_d") %>%
  rename(littersize = "x15_1_litter_size") %>%
  rename(litterpyr = "x16_1_litters_per_year") %>%
  rename(longev = "x17_1_max_longevity_m") %>%
  rename(neonatemass = "x5_3_neonate_body_mass_g")%>%
  rename(sexmat = "x23_1_sexual_maturity_age_d")%>%
  mutate(longev *30)%>%
  rename(longevd = "longev * 30")

pi2 <- pi %>%
  select(species, placenta_invasiveness, X.4, interdigitation_placenta, X.5)
# Un-reverse score the numeric ordinal variables
pi2$invade = pi2$X.4
pi2$invade[pi2$invade == "1"] <- "4" # used 4 to get these values out of the way for renaming "3's" as "1's"
pi2$invade[pi2$invade == "3"] <- "1"
pi2$invade[pi2$invade == "4"] <- "3"

pi2$digit = pi2$X.5
pi2$digit[pi2$digit == "1"] <- "4" # same here
pi2$digit[pi2$digit == "3"] <- "1"
pi2$digit[pi2$digit == "4"] <- "3"
#select variables with correctly scored numeric ordinal variables
pi3 <- pi2 %>%
  select(species, placenta_invasiveness, invade, interdigitation_placenta, digit)

np2 <- np %>%
  select(KeepWildRecords.false, X.5, X.7, X.9, X.10, X.11) %>%
  rename(species = "KeepWildRecords.false") %>%
  rename(tot.neo = "X.5") %>%
  rename(prev.neo = "X.7") %>%
  rename(tot.mal = "X.9") %>%
  rename(prev.mal = "X.10") %>%
  rename(mal.trans = "X.11")

# Life history + placenta data
combo1 <- merge(lh2, pi3, by= "species")

anticombo1 <- anti_join(lh2, pi3, by = "species") # list of animals we have LH data for, but not placenta data
# Filter out missing values
combo2 <- combo1 %>%
  filter(!agefb==-999.00) %>% # -999.0 = missing data
  filter(!adultmass==-999.00)%>%
  filter(!gestation==-999.00) %>%
  filter(!gestation==-999.00) %>%
  filter(!interv==-999.00) %>%
  filter(!litterpyr==-999.00) %>%
  filter(!longev==-999.00) %>%
  filter(!sexmat==-999.00)%>%
  filter(!neonatemass==-999.00)
# Only 53 species with complete LH + placenta data!

# Life history + placental + cancer data
megaset <- merge(combo1, np2, by= "species")

antimega <- anti_join(combo1, np2, by= "species") # list of animals we have LH + placenta data for, but not cancer data

# Life history + cancer data
lifecancer <- merge(lh2, np2, by = "species")
lifecancer$tot.neo <- as.numeric(lifecancer$tot.neo) # set rates as numeric
lifecancer$prev.neo <- as.numeric(lifecancer$prev.neo)
lifecancer$tot.mal <- as.numeric(lifecancer$tot.mal)
lifecancer$prev.mal <- as.numeric(lifecancer$prev.mal)
lifecancer$mal.trans <- as.numeric(lifecancer$mal.trans)

lifecancer2 <- lifecancer %>%
  mutate_all(funs(str_replace(.," ","_"))) %>%
   filter(!gestation==-999.00)
# 143 species with complete LH + cancer data

# Placentas + Cancer data
placentacancer <- merge(pi3, np2, by = "species")

placentacancer1 <- placentacancer[-c(132),] # merge threw column labels in the data set, remove this row!
placentacancer1$invade <- as.numeric(placentacancer1$invade) # set rates as numeric
placentacancer1$digit <- as.numeric(placentacancer1$digit)
placentacancer1$tot.neo <- as.numeric(placentacancer1$tot.neo)
placentacancer1$prev.neo <- as.numeric(placentacancer1$prev.neo)
placentacancer1$tot.mal <- as.numeric(placentacancer1$tot.mal)
placentacancer1$prev.mal <- as.numeric(placentacancer1$prev.mal)
placentacancer1$mal.trans <- as.numeric(placentacancer1$mal.trans)
#Rename so graphs are prettier :)
placentacancer1$placenta_invasiveness[placentacancer1$placenta_invasiveness == "epi"] <- "Epitheliochorial"
placentacancer1$placenta_invasiveness[placentacancer1$placenta_invasiveness == "endo"] <- "Endotheliochorial"
placentacancer1$placenta_invasiveness[placentacancer1$placenta_invasiveness == "hemo"] <- "Hemochorial"

placentacancer2 <- placentacancer1 %>%
  mutate_all(funs(str_replace(.," ","_"))) # replace all spaces in species names with underscores to be consistent with phylogeny data
```

# Phylogeny wrangling
Comments on this section would be pretty long to explain in each step. In essence, this code was computed to ensure there is complete data for the PGLS models. I'd recommend running this code line by line & taking a look at the object/output to visualize and make sense of what is happening in each step. Feel free to reach out if you'd like further explanation!
```{r}
# Phylogeny - Mammals + Placentas + Cancer (no Life History)
species_placcancer <- placentacancer2 %>%
  dplyr::select("species")

mammals_list <- as.phylo(mammals)

specmammal <- as.data.frame(mammals_list$tip.label)%>%
  rename("species"="mammals_list$tip.label")

specieslist <- merge(specmammal, species_placcancer, by.x = "species", by.y = "species")

specieslist_v <- as.vector(specieslist$species)

specieslist_k <- keep.tip(mammals, specieslist_v)

write.csv(specieslist_v, "specieslist_v.csv", row.names = FALSE, col.names = TRUE)

specieskept <- read.csv("specieslist_v.csv")

mammallist <- merge(specieskept, placentacancer2, by.x = "x", by.y = "species")%>%
  rename("species"=x)

mammallist1 <- mammallist
mammallist1[is.na(mammallist1)] <- 0

mammal.table <- as_tibble(mammals)
as.phylo(mammals)


#Phylogeny - Mammals + Life History + Cancer
species_lifecancer <- lifecancer2 %>%
  dplyr::select("species")

specmammal2 <- as.data.frame(mammals_list$tip.label)%>%
  rename("species"="mammals_list$tip.label")

specieslist2 <- merge(specmammal2, species_lifecancer, by.x = "species", by.y = "species")

specieslist_v2 <- as.vector(specieslist2$species)

specieslist_k2 <- keep.tip(mammals, specieslist_v2)

write.csv(specieslist_v2, "specieslist_v2.csv", row.names = FALSE, col.names = TRUE)

specieskept2 <- read.csv("specieslist_v2.csv")

mammallist2 <- merge(specieskept2, lifecancer2, by.x = "x", by.y = "species")%>%
  rename("species"=x)

mammallist3 <- mammallist2
mammallist3[is.na(mammallist3)] <- 0

mammal.table <- as_tibble(mammals)
as.phylo(mammals)

```
Okay all the wrangling is finally done...

One of my hypotheses for life history traits that may better explain cancer vulnerability rather than placental morphology was maximum offspring output. Basically the quantity over quality concept of life history/cancer discussions to the absolute extreme. Here I went through our LH data & computed for maximum offspring output or "maximum babies per lifetime". This data set is used later.

# Life History computations
```{r}
# Computing for births per lifetime:
#   Lifespan - age at sexual maturation (assumes no post-reproductive period) = Reproductive lifespan
combo2.1 <- combo2 %>%
  mutate(longevd - sexmat)%>%
  rename(reprolife = "longevd - sexmat") 

#   Gestation length + interbirth interval = Time per litter
combo2.2 <- combo2.1 %>%
 mutate(gestation + interv) %>%
 rename(timeprlit = "gestation + interv")

#   Reproductive lifespan / time per litter = Birthing potential
combo2.3 <- combo2.2 %>%
  mutate(reprolife / timeprlit) %>%
  rename(birthpotent = "reprolife/timeprlit")

# Computing for offspring output:
#   Maximum births per lifetime * Avg litter size
combo2.4 <- combo2.3 %>%
  mutate(birthpotent * littersize) %>%
  rename(maxoffspring = "birthpotent * littersize")

# throw it together in a master data set to use later
megaset5 <- merge(combo2.4, placentacancer1, by = "species")
```

Now we should have all the data sets we need & it's time to start doing analyses. The first step is to look at ANOVAs. If you're not familiar, ANOVAs test for variance in means across groups. The ANOVAs don't tell us a whole lot on their own, so I assigned each ANOVA to its own object for "best fit" modeling later.

# Placental Invasion x Cancer ANOVAs
```{r}
# Malignant transformation as a function of placental invasion
pmtanova <- aov(mal.trans ~ invade, data = placentacancer1)

summary(pmtanova) # 22 N/As

# Malignancy prevalence as a function of placental invasion
pmpanova <- aov(prev.mal ~ invade, data = placentacancer1)

summary(pmpanova)

# Neoplasia prevalence as a function of placental invasion
pnpanova <- aov(prev.neo ~ invade, data = placentacancer1)

summary(pnpanova)

```

# Interdigitation x Cancer ANOVAs
```{r}
# Malignant transformation as a function of interdigitation
imtanova <- aov(mal.trans ~ digit, data = placentacancer1)

summary(imtanova)

# Malignancy prevalence as a function of interdigitation
impanova <- aov(prev.mal ~ digit, data = placentacancer1)

summary(impanova)

# Neoplasia prevalence as a function of interdigitation
inpanova <- aov(prev.neo ~ digit, data = placentacancer1)

summary(inpanova)
```

# Invasion + Interdigitation x Cancer ANOVAs (additive testing)
```{r}
# Malignant transformation as a function of invasion + interdigitation
pimtanova <- aov(mal.trans ~ invade + digit, data = placentacancer1)

summary(pimtanova)

# Malignancy prevalence as a function of invasion + interdigitation
pimpanova <- aov(prev.mal ~ invade + digit, data = placentacancer1)

summary(pimpanova)

# Neoplasia prevalence as a function of invasion + interdigitation
pinpanova <- aov(prev.neo ~ invade + digit, data = placentacancer1)

summary(pinpanova)
```

Invasion * Interdigitation x Cancer ANOVAs (interaction testing)
```{r}
# Malignant transformation as a function of invasion * interdigitation
pimtanova1 <- aov(mal.trans ~ invade * digit, data = placentacancer1)

summary(pimtanova1) #22 N/As

# Malignancy prevalence as a function of invasion * interdigitation
pimpanova1 <- aov(prev.mal ~ invade * digit, data = placentacancer1)

summary(pimpanova1)

# Neoplasia prevalence as a function of invasion * interdigitation
pinpanova1 <- aov(prev.neo ~ invade * digit, data = placentacancer1)

summary(pinpanova1)
```

Okay! So now we have tested for variance & have a bunch of numbers that aren't super telling. Now we will test the ANOVAs against one another through Akaike information criterion (AICs) to determine which ANOVA model is th "best fit".

# Invasion AICs
```{r}
# Malignant Transformation
mtset <- list(pmtanova, pimtanova, pimtanova1)
mtnames <- c("pmtanova", "pimtanova", "pimtanova1")

aictab(mtset, modnames = mtnames)
        #"pimtanova1",interaction model is the best fit.

par(mfrow=c(2,2))
plot(pimtanova1)
par(mfrow=c(1,1))



# Malignancy Prevalence
mpset <- list(pmpanova, pimpanova, pimpanova1)
mpnames <- c("pmpanova", "pimpanova", "pimpanova1")

aictab(mpset, modnames = mpnames)
      #"pimpanova", additive model is the best fit.

par(mfrow=c(2,2))
plot(pimpanova)
par(mfrow=c(1,1))


# Neoplasia Prevalence
npset <- list(pnpanova, pinpanova, pinpanova1)
npnames <- c("pnpanova", "pinpanova", "pinpanova1")

aictab(npset, modnames = npnames)
      #"pnpanova", solo model is the best fit

par(mfrow=c(2,2))
plot(pnpanova)
par(mfrow=c(1,1))

```

# Interdigitation AICs
```{r}
# Malignant Transformation
mtsetd <- list(imtanova, pimtanova, pimtanova1)
mtnamesd <- c("imtanova", "pimtanova", "pimtanova1")

aictab(mtsetd, modnames = mtnamesd)
      #"imtanova", solo model is the best fit, interactive a close 2nd

par(mfrow=c(2,2))
plot(imtanova)
par(mfrow=c(1,1))

par(mfrow=c(2,2))
plot(pimtanova1)
par(mfrow=c(1,1))


# Malignancy Prevalence
mpsetd <- list(impanova, pimpanova, pimpanova1)
mpnamesd <- c("impanova", "pimpanova", "pimpanova1")

aictab(mpsetd, modnames = mpnamesd)
      #"impanova", solo model is the best fit

par(mfrow=c(2,2))
plot(impanova)
par(mfrow=c(1,1))


# Neoplasia Prevalence
npsetd <- list(inpanova, pinpanova, pinpanova1)
npnamesd <- c("inpanova", "pinpanova", "pinpanova1")

aictab(npsetd, modnames = npnamesd)
      #"inpanova", solo model is the best fit

par(mfrow=c(2,2))
plot(inpanova)
par(mfrow=c(1,1))
```

I don't know why my Tukeys stopped working, looking into this but ignore for now.

# Tukey's HSD 
```{r}
# Malignant Transfromation as a function of the interaction of Placental Invasion & Interdigitation
#tukey.pimt <- TukeyHSD(pimtanova1)
#plot(tukey.pimts)

# Malignant Transformation as a function of Placental Interdigitation
#tukey.imt <- TukeyHSD(imtanova)
#plot(tukey.imt)




#Malignancy Prevalence as a function of added Invasion + Interdigitation
#tukey.pimp <- TukeyHSD(pimpanova)
#plot(tukey.pimp)

# Malignancy Prevalence as a function of Interdigitation
#tukey.imp <- TukeyHSD(impanova)
#plot(tukey.imp)




# Neoplasia Prevalence as a function of Placental Invasion
#tukey.pnp <- TukeyHSD(pnpanova)
#plot(tukey.pnp)

# Neoplasia Prevalence as a function of Interdigitation
#tukey.inp <- TukeyHSD(inpanova)
#plot(tukey.inp)
```

Now we'll do the same thing must use the maximum offspring output variable we created earlier.

# Birth Potential x Cancer ANOVAs (nothing significant)
```{r}
# Malignant transformation as a function of birth potential
bpmtanova <- aov(mal.trans ~ birthpotent, data = megaset5)

summary(bpmtanova)

# Malignancy prevalence as a function of birth potential
bpmpanova <- aov(prev.mal ~ birthpotent, data = megaset5)

summary(bpmpanova)

# Neoplasia prevalence as a function of birth potential
bpnpanova <- aov(prev.neo ~ birthpotent, data = megaset5)

summary(bpnpanova)

```

# Maximum Offspring x Cancer ANOVAs (nothing significant)
```{r}
# Malignant transformation as a function of maximum offspring
momtanova <- aov(mal.trans ~ maxoffspring, data = megaset5)

summary(momtanova)

# Malignancy prevalence as a function of maximum offspring
mompanova <- aov(prev.mal ~ maxoffspring, data = megaset5)

summary(mompanova)

# Neoplasia prevalence as a function of maximum offspring
monpanova <- aov(prev.neo ~ maxoffspring, data = megaset5)

summary(monpanova)
```

Color Palette
```{r}
earth <- c("#61bc36", "#365a54", "#444444")

orange <- c("#FF8459", "#CC530A", "#94280A")

blue <- c("#A8D5C0", "#599BA1", "#1E4E64")
```

Life history graphs
```{r}
# Lifespan x Time per litter x Placental invasion
combo2.4 %>%
  ggplot(.,aes(x=longevd, y=timeprlit, color= invade))+
  geom_point()+
  geom_smooth(method = "lm", se= FALSE)+
  geom_label_repel(aes(label = species))


# Lifespan x Time per litter x Interdigitation
combo2.4 %>%
  ggplot(.,aes(x=longevd, y=timeprlit, color= digit))+
  geom_point()+
  geom_smooth(method = "lm", se= FALSE)+
  geom_label_repel(aes(label = species))

# Maximum birth potential x Placental Invasion
combo2.4 %>%
  ggplot(., aes(y= birthpotent, x= placenta_invasiveness))+
  geom_boxplot()

# Gestation x Invasiveness
combo2$placenta_invasiveness[combo2$placenta_invasiveness == "epi"] <- "Epitheliochorial"
combo2$placenta_invasiveness[combo2$placenta_invasiveness == "endo"] <- "Endotheliochorial"
combo2$placenta_invasiveness[combo2$placenta_invasiveness == "hemo"] <- "Hemochorial"
combo2$placenta_invasiveness <- factor(combo2$placenta_invasiveness,levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))
combo2 %>%
  ggplot(., aes(x= placenta_invasiveness, y= gestation))+
  geom_boxplot()

# Gestation x Interdigitation
combo2$interdigitation_placenta <- factor(combo2$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))

ggplot(combo2, aes(x= interdigitation_placenta, y= gestation, fill=interdigitation_placenta))+
  geom_boxplot()+
  ggtitle("Gestation Length by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Gestation Length (Days)")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")

# Neonate Mass x Interdigitation
combo2$interdigitation_placenta <- factor(combo2$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))
ggplot(combo2, aes(x= interdigitation_placenta, y= neonatemass, fill=interdigitation_placenta))+
  geom_boxplot()+
  ggtitle("Neonate Mass by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Neonate Mass (g)")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")

# Adult Mass x Interdigitation
combo2$interdigitation_placenta <- factor(combo2$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))
ggplot(combo2, aes(x= interdigitation_placenta, y= adultmass, fill=interdigitation_placenta))+
  geom_boxplot()+
  ylim(c(0,750000))+
  ggtitle("Adult Mass by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Adult Mass (g)")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")

# Birth Potential x Interdigitation
combo2.4$interdigitation_placenta <- factor(combo2$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))
ggplot(combo2.4, aes(x= interdigitation_placenta, y= birthpotent, fill=interdigitation_placenta))+
  geom_boxplot()+
  ggtitle("Birth Potential by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Max # of births per lifetime")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")

# Max Offspring Output x Interdigitation
combo2.4$interdigitation_placenta <- factor(combo2$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))
ggplot(combo2.4, aes(x= interdigitation_placenta, y= maxoffspring, fill=interdigitation_placenta))+
  geom_boxplot()+
  ylim(c(0,150))+
  ggtitle("Maximum Offspring Output by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Max # of offspring per lifetime")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")

# Max Longevity x Interdigitation
combo2$interdigitation_placenta <- factor(combo2$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))
ggplot(combo2, aes(x= interdigitation_placenta, y= longevd, fill=interdigitation_placenta))+
  geom_boxplot()+
  ggtitle("Max Longevity by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Max Longevity (Days)")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")

# Age of Sexual Maturity x Interdigitation
combo2$interdigitation_placenta <- factor(combo2$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))
ggplot(combo2, aes(x= interdigitation_placenta, y= sexmat, fill=interdigitation_placenta))+
  geom_boxplot()+
  ggtitle("Age at Sexual Maturity by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Age at Sexual Maturity (Days)")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")

```

Placentas & Cancer Graphs
```{r}

# Placental invasion x neoplasia prevalence
placentacancer1$placenta_invasiveness <- factor(placentacancer1$placenta_invasiveness,levels = c("Epitheliochorial", "Endotheliochorial", "Hemochorial"))

placentacancer1 %>%
  ggplot(.)+
  geom_violin(aes(x = placenta_invasiveness, y = prev.neo, fill=placenta_invasiveness))+
  ggtitle("Neoplasia Prevalence by Invasiveness")+
  xlab("Placenta Invasiveness")+
  ylab("Neoplasia %")+
  theme_pander()+
  scale_fill_manual(values = orange)+
  theme(legend.position = "none")


# Interdigitation x neoplasia prevalence
placentacancer1$interdigitation_placenta <- factor(placentacancer1$interdigitation_placenta,levels = c("Villous", "Trabecular", "Labyrinthine"))


placentacancer1 %>%
  ggplot(.)+
  geom_violin(aes(x= interdigitation_placenta, y= prev.neo, fill=interdigitation_placenta))+
  ggtitle("Neoplasia Prevalence by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Neoplasia %")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")

# Placental invasion x malignancy prevalence

placentacancer1 %>%
  ggplot(.)+
  geom_violin(aes(x = placenta_invasiveness, y = prev.mal, fill=placenta_invasiveness))+
  ggtitle("Malignancy Prevalence by Invasiveness")+
  xlab("Placenta Invasiveness")+
  ylab("Malignant %")+
  theme_pander()+
  scale_fill_manual(values = orange)+
  theme(legend.position = "none")


# Interdigitation x malignancy prevalence

placentacancer1 %>%
  ggplot(.)+
  geom_violin(aes(x= interdigitation_placenta, y= prev.mal, fill=interdigitation_placenta))+
  ggtitle("Malignancy Prevalence by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Malignant %")+
  theme_pander()+
  scale_fill_manual(values = blue)+
  theme(legend.position = "none")


# Placental invasion x malignant transformation rate

placentacancer1 %>%
  ggplot(.)+
  geom_violin(aes(x= placenta_invasiveness, y= mal.trans, fill=placenta_invasiveness))+
  ggtitle("Rate of Malignant Transformation by Invasiveness")+
  xlab("Placenta Invasiveness")+
  ylab("Rate of Malignant Transformation")+
  theme_pander(lp= "none")+
  scale_fill_manual(values = orange)+
  theme(legend.position = "none")


# Interdigitation x malignant transformation rate
placentacancer1 %>%
  ggplot(.)+
  geom_violin(aes(x= interdigitation_placenta, y= mal.trans, fill=interdigitation_placenta))+
  ggtitle("Rate of Malignant Transformation by Interdigitation")+
  xlab("Interdigitation Class")+
  ylab("Rate of Malignant Transformation")+
  theme_pander(lp="none")+
  scale_fill_manual(values = blue)





```

Birth Potential Graphs
```{r}
# Birth potential x interdigitation
combo2.4 %>%
  ggplot(.) +
  geom_boxplot(aes(x= interdigitation_placenta, y= birthpotent))

# Maximum offspring output x placental invasion
combo2.4 %>%
  ggplot(.) +
  geom_boxplot(aes(x= placenta_invasiveness, y= birthpotent))

# Maximum offspring output x neoplasia prevalence
megaset5 %>%
  ggplot(.)+
  geom_point(aes(x= birthpotent, y= prev.neo))

# Maximum offspring output x malignancy prevalence
megaset5 %>%
  ggplot(.)+
  geom_point(aes(x= birthpotent, y= prev.mal))

# Maximum offspring output x malignant transformation rate
megaset5 %>%
  ggplot(.)+
  geom_point(aes(x= birthpotent, y= mal.trans))
```

Placental Invasion x Cancer Correlation Computations
```{r}
# Placental Invasion x Malignant Transformation

invmal1 <- lm(placentacancer1$invade ~ placentacancer1$mal.trans)
summary(invmal1)

# Placental Invasion x Neoplasia Prevalence

invneo1 <- lm(placentacancer1$invade ~ placentacancer1$prev.neo)
summary(invneo1)

# Placental Invasion x Malignancy Prevalence

invmalprev1 <- lm(placentacancer1$invade ~ placentacancer1$prev.mal)
summary(invmalprev1)

```

Interdigitation Correlation x Cancer Computations
```{r}
# Interdigitation x Malignant Transformation

digmal1 <- lm(placentacancer1$digit ~ placentacancer1$mal.trans)
summary(digmal1)

# Interdigitation x Neoplasia Prevalence

digneo1 <- lm(placentacancer1$digit ~ placentacancer1$prev.neo)
summary(digneo1)

# Interdigitation x Malignancy Prevalence

digmalprev1 <- lm(placentacancer1$digit ~ placentacancer1$prev.mal)
summary(digmalprev1)

```

Life History x Cancer Correlation Computations
```{r}
# Lifespan x Malignant Transformation

longevdmal1 <- lm(lifecancer$longevd ~ lifecancer$mal.trans)
summary(longevdmal1)

# Lifespan x Neoplasia Prevalence

longevdneo1 <- lm(lifecancer$longevd ~ lifecancer$prev.neo)
summary(longevdneo1)

# Lifespan x Malignancy Prevalence

longevdmalprev1 <- lm(lifecancer$longevd ~ lifecancer$prev.mal)
summary(longevdmalprev1)




# Birth Potential x Malignant Transformation
birthpotmal <- lm(megaset5$birthpotent ~ megaset5$mal.trans)
summary(birthpotmal)

# Birth Potential x Neoplasia Prevalence
birthpotneo <- lm(megaset5$birthpotent ~ megaset5$prev.neo)
summary(birthpotneo)

# Birth Potential x Malignancy Prevalence
birthpotmalprev <- lm(megaset5$birthpotent ~ megaset5$prev.mal)
summary(birthpotmalprev)




# Max Offspring x Malignant Transformation
maxoffmal <- lm(megaset5$maxoffspring ~ megaset5$mal.trans)
summary(maxoffmal)

# Max Offspring x Neoplasia Prevalence
maxoffneo <- lm(megaset5$maxoffspring ~ megaset5$prev.neo)
summary(maxoffneo)

# Max Offspring x Malignancy Prevalence
maxoffmalprev <- lm(megaset5$maxoffspring ~ megaset5$prev.mal)
summary(maxoffmalprev)

```

PGLS - Placentas + Cancer (Brownian)
```{r}

# Malignant Transformation x Interdigitation (browninan)
malig <- mammallist1[, "mal.trans"]
interdig <- mammallist1[, "digit"]

malig1 <- as.numeric(malig)
interdig1 <- as.numeric(interdig)

pgls_interdig <- gls(malig1 ~ interdig1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_interdig)
coef(pgls_interdig)


anova.pgls.mti <- anova(pgls_interdig)


# Neoplasia Prevalence x Interdigitation (Brownian)
neop <- mammallist1[,"prev.neo"]

neop1 <- as.numeric(neop)

pgls_neointerdig <- gls(neop1 ~ interdig1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_neointerdig)
coef(pgls_neointerdig)
pgls_neointerdig1 <- anova(pgls_neointerdig)
summary(pgls_neointerdig1)

# Malignancy Prevalence by Interdigitation (Brownian)
mal <- mammallist1[,"prev.mal"]
mal1<- as.numeric(mal)

pgls_malinterdig <- gls(mal1 ~ interdig1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_malinterdig)

# Malignant Transformation x Invasion (Brownian)
invasion <- mammallist1[, "invade"]
invasion1 <- as.numeric(invasion)

pgls_invasion <- gls(malig1 ~ invasion1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_invasion)

pgls_invasion1 <- anova(pgls_invasion)
summary(pgls_invasion1)

invasion1.df <- as.data.frame(pgls_invasion1)



# Neoplasia Prevalence x Invasion (Brownian)
pgls_npinvade <- gls(neop1 ~ invasion1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_npinvade)

npinvads.anova <- anova(pgls_npinvade)
summary(npinvads.anova)

pgls_malinvade <- gls(mal1 ~ invasion1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_malinvade)

# Neoplasia Prevalence x both (interactive)
pgls_npboth <- gls(neop1 ~ invasion1*interdig1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_npboth)

# Malignancy Prevalence x both (interactive)
pgls_mpboth <- gls(mal1 ~ invasion1*interdig1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_mpboth)

#Malignant Transformaiton by both (interactive)
pgls_mtboth <- gls(malig1 ~ interdig1*invasion1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_mtboth)
coef(pgls_mtboth)

# Neoplasia Prevalence x both (additive)
pgls_npboth.add <- gls(neop1 ~ invasion1+interdig1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_npboth.add)


#Malignant Transformaiton by both (additive)
pgls_mtboth.add <- gls(malig1 ~ interdig1+invasion1, correlation = corBrownian(phy = mammals_list), data = mammallist, method = "ML")
summary(pgls_mtboth.add)
coef(pgls_mtboth)





```

PGLS - Placentas + Cancer (Ornstein-Uhlenbeck)
```{r}
# Malignant Transformation x Interdigitation (Ornstein-Uhlenbeck)
ou.pgls.interdig <-corMartins(1,phy=mammals_list)
ou.interdig<-gls(malig1~interdig1,correlation=ou.pgls.interdig,data=mammallist)
summary(ou.interdig)

# Neoplasia Prevalence x Interdigitation (Ornstein-Uhlenbeck)
ou.pgls.neointerdig <-corMartins(1,phy=mammals_list)
ou.neointerdig<-gls(neop1~interdig1,correlation=ou.pgls.neointerdig,data=mammallist)
summary(ou.neointerdig)

# Malignancy Prevalence x Interdigitation (Ornstein-Uhlenbeck)
ou.pgls.malinterdig <-corMartins(1,phy=mammals_list)
ou.malinterdig<-gls(mal1~interdig1,correlation=ou.pgls.neointerdig,data=mammallist)
summary(ou.malinterdig)

# Malignant Transformation x Invasion (Ornstein-Uhlenbeck)
ou.pgls.invasion <-corMartins(1,phy=mammals_list)
ou.invasion<-gls(malig1~invasion1,correlation=ou.pgls.invasion,data=mammallist)
summary(ou.invasion)

# Neoplasia Prevalence x Invasion (Ornstein-Uhlenbeck)
ou.pgls.neoinvasion <-corMartins(1,phy=mammals_list)
ou.neoinvasion1<-gls(neop1~invasion1,correlation=ou.pgls.neoinvasion,data=mammallist)
summary(ou.neoinvasion1)

#Malignancy Prevalence x Invasion (Ornstein-Uhlenbeck)
ou.pgls.malinvade <-corMartins(1,phy=mammals_list)
ou.malinvade<-gls(mal1~invasion1,correlation=ou.pgls.malinvade,data=mammallist)
summary(ou.malinvade)

# Neoplasia Prevalence x both (interactive)
ou.pgls_npboth <- corMartins(1,phy=mammals_list)
ou.npboth <- gls(neop1 ~ invasion1*interdig1, correlation = ou.pgls_npboth, data = mammallist)
summary(ou.npboth)


#Malignant Transformaiton by both (interactive)
ou.pgls_mtboth <- corMartins(1,phy=mammals_list)
ou.mtboth <- gls(malig1 ~ interdig1*invasion1, correlation = ou.pgls_mtboth, data = mammallist)
summary(ou.mtboth)

#Malignancy Prevalence by both (interactive)
ou.pgls.mpboth <-corMartins(1,phy=mammals_list)
ou.mpboth<-gls(mal1~interdig1*invasion1,correlation=ou.pgls.mpboth,data=mammallist)
summary(ou.mpboth)

# Neoplasia Prevalence x both (additive)
ou.pgls_npboth.add <- corMartins(1,phy = mammals_list)
ou.npboth.add <- gls(neop1 ~ invasion1+interdig1, correlation = ou.pgls_npboth.add, data = mammallist)
summary(ou.npboth.add)


#Malignant Transformaiton by both (additive)
ou.pgls_mtboth.add <- corMartins(1,phy= mammals_list)
ou.mtboth.add <- gls(malig1 ~ interdig1+invasion1, correlation = ou.pgls_mtboth.add, data = mammallist)
summary(ou.mtboth.add)
```

PGLS - Life History + Cancer
```{r}
# Gestation Length x Neoplasia Prevalence
gest <- mammallist3[, "gestation"]
gest1 <- as.numeric(gest)
 
neop2 <- mammallist3[, "prev.neo"]
neop3 <- as.numeric(neop2)

pgls_gestnp <- gls(neop3 ~ gest1, correlation = corBrownian(phy = mammals_list), data = mammallist3, method = "ML")

summary(pgls_gestnp)
coef(pgls_gestnp)

# Gestation Length x Malignant Transformation
mammallist4 <- mammallist3 %>%
  na.omit(mammallist3)

gest2 <- mammallist4[, "gestation"]
gest3 <- as.numeric(gest2)

malig2 <- mammallist4[, "mal.trans"]
malig3 <- as.numeric(malig2) 

pgls_gestmt <- gls(malig3 ~ gest3, correlation = corBrownian(phy = mammals_list), data = mammallist4, method = "ML")

summary(pgls_gestmt)
coef(pgls_gestmt)

```






FGQ and LGQ x Cancer PGLS
```{r}
gest_cancer <- merge(gestation_neonate, np2, by = "species")
lit_cancer <- merge(gestator, np2, by = "species")



#load in phylogenic information, combine with placenta and lh info
mammal_phy <- ggtree(tree$mammalST_bestDates)
#%<+% phylo_df
#makes into phylo class from ggtree
mammal_phy<-as.phylo(mammal_phy)
#Make data set for the phylogenic correction that has the placenta variables, neonate body size, gestation length, and GQ/LGQ 
gestation_neonate_phylo1<-gest_cancer %>% 
  dplyr::select(species,prev.neo,prev.mal, mal.trans, gq)
# Next step is to remove mammals that are not represented in placenta df. We do this by creating a list of ones that dont match and then removing them.
species_cancer<-gestation_neonate_phylo1 %>% 
  dplyr::select("species")
species_phylo<-as.data.frame(mammal_phy$tip.label) %>% 
  mutate_all(funs(str_replace(.,"_"," ")))
species_kept<-merge(species_cancer,species_phylo,by.y="mammal_phy$tip.label",by.x = "species") %>% 
  mutate_all(funs(str_replace(.," ","_")))
species_kept_v<-as.vector(species_kept$species)
small_mammal_k<-keep.tip(mammal_phy,species_kept_v)
write.csv(species_kept_v,"species_kept_v.csv",row.names = FALSE,col.names = TRUE)
species_kept_p<-read.csv("species_kept_v.csv",header=TRUE)
#clean placenta data so that it fits into the phylogeny information
gestation_neonate_phylo2<-gestation_neonate_phylo1 %>% 
  mutate_all(funs(str_replace(.," ","_")))
gestation_neonate_phylo2$prev.neo<-as.numeric(gestation_neonate_phylo2$prev.neo)
gestation_neonate_phylo2$prev.mal<-as.numeric(gestation_neonate_phylo2$prev.mal)
gestation_neonate_phylo2$mal.trans<-as.numeric(gestation_neonate_phylo2$mal.trans)
gestation_neonate_phylo2$gq<-as.numeric(gestation_neonate_phylo2$gq)

#make comparative DF
#comp_phylo2<-comparative.data(mammal_phy,gestation_neonate_phylo2, "species", vcv=FALSE, vcv.dim=2, na.omit=TRUE, 
                 #force.root=FALSE, warn.dropped=FALSE, scope=NULL)

#PGLS
gestation_neonate_phylo3 <- gestation_neonate_phylo2 %>%
  drop_na()

  #FGQ x Malignant Transformation
fgq_maltrans <- gls(gq ~ mal.trans, correlation = corBrownian(phy = mammal_phy), data = gestation_neonate_phylo3, method = "ML", )
summary(fgq_maltrans) # P value 0.2309
AIC(fgq_maltrans) #512.8376

  #FGQ x Malignancy Prevalence
fgq_malprev <- gls(gq ~ prev.mal, correlation = corBrownian(phy = mammal_phy), data = gestation_neonate_phylo3, method = "ML", )
summary(fgq_malprev) # P value 0.9213
AIC(fgq_malprev) #514.3031

  #FGQ x Neopasia Prevalence
fgq_neoprev <- gls(gq ~ prev.neo, correlation = corBrownian(phy = mammal_phy), data = gestation_neonate_phylo3, method = "ML", )
summary(fgq_neoprev) # P value 0.7297
AIC(fgq_neoprev) #514.1904

```


